IFC Inside: A General Approach to Retrofitting Languages with Dynamic Information Flow Control

### 2014-07-08 paper/retrofit.tex
Minor changes to retrofit.tex

Information flow control systems operate by associating data with \emph{labels},
which specify whether or not data tagged with one label |l1| can flow to another
label |l2| (written as |canFlowTo2 l1 l2|).  These labels encode the desired
security policy (for example, confidential information should not flow to
a public channel), and the work of specifying the semantics of an information
flow language involves demonstrating that impermissible flows cannot happen,
a property called \emph{noninterference}.

@@

Why is this setting more interesting than ``traditional
coarse-grained information flow control system on isolated
threads?''  First, although completely separate tasks
are semantically easy to reason about, real-world implementations often blur the
lines between tasks in the name of efficiency.  Characterizing what optimizations
are permissible is a problem we address in Section~\ref{sec:concrete}, and
it is quite subtle, since removing transitions from the operational semantics
of a language can break noninterference.  Second, even though the underlying
semantic model involves communicating tasks, these tasks can be coordinated
together in ways that simulate traditional features in monolithic languages.
In fact, it is a very useful design tool for discovering what variants of the
features are permissible and which are not.

A very simple example is a low security task sending a high security
task a mutable reference:
\begin{align*}
    & |let i = TI (sandbox (brecvf x _ (IT (!(TI x)))))|\\
    & |in TI (send (IT i) getLabel (IT (ref true)))|
\end{align*}

For now, ignore |TI| and |IT|: roughly, this code creates a new
|sandbox|ed thread $i$ which waits (\textbf{blockingRecv}) for a
message, and then \textbf{send}s the thread a mutable reference to (|ref
true|).  If this operation actually shared the mutable cell, this cell
could be used to violate information flow control.  When working with
isolated tasks which each have their own heaps, the isolation is obviously true:
the sandboxed task receives a dangling pointer.

While this is a very simple example, we can still ask: given the
obviously safe formulation with separate heaps, can this fact be useful
for showing the safety of a semantics which only has a single heap
(Yes: see Section~\ref{sec:single-heap})?  Furthermore, the ``raw''
semantics violates expectations about how inter-thread communication
works on a shared heap: it strongly suggests, for example, that which
types of data may be passed between tasks should be restricted (for
example, in JavaScript, one might only allow JSON objects to be passed
between tasks).

@@

When considering tasks, we will instead say that the allocation of a new
stack frame with a raised label corresponds to the creation of a new
isolated task; a sample encoding using our language is shown in
Figure~\ref{fig:toLabeled}.
However, this is not enough: in a stack discipline, one generally
expects stack entries to be processed in order, rather than in parallel
as would be suggested by isolated tasks.  The sequential processing of
the task queue corresponds to a specific \emph{scheduling policy} of
these threads.  The recasting of the problem to one of scheduling
highlights the essential difference between termination-insensitive and
termination-sensitive non-interference.

>>>>

Information flow control systems operate by associating data with \emph{labels},
and specifying whether or not data tagged with one label |l1| can flow
to another
label |l2| (written as |canFlowTo2 l1 l2|).  These labels encode the desired
security policy (for example, confidential information should not flow to
a public channel), and the work of specifying the semantics of an information
flow language involves demonstrating that impermissible flows cannot happen,
a property called \emph{non-interference}.

@@

Why is this setting more interesting than ``traditional
coarse-grained information flow control system on isolated
threads?''  First, although completely separate tasks
are semantically easy to reason about, real-world implementations often blur the
lines between tasks in the name of efficiency.  Characterizing what optimizations
are permissible is a problem we address in Section~\ref{sec:concrete}, and
it is quite subtle, since removing transitions from the operational semantics
of a language can in general break non-interference.
Second, even though the underlying
semantic model involves communicating tasks, these tasks can be coordinated
together in ways that simulate traditional features in monolithic languages.
In fact, it is a very useful design tool for discovering what variants of the
features are permissible and which are not.

A very simple example is a low security task sending a high security
task a mutable reference:
\begin{align*}
    & |let i = TI (sandbox (brecvf x _ (IT (!(TI x)))))|\\
    & |in TI (send (IT i) getLabel (IT (ref true)))|
\end{align*}
For now, ignore the tags |TI| and |IT|: roughly, this code creates a new
|sandbox|ed task with identifier $i$ which waits (\textbf{blockingRecv}) for a
message, and then \textbf{send}s the task a mutable reference to |(ref
true)|.  If this operation actually shared the mutable cell, this cell
could be used to violate information flow control, by leaking
information from the high security task to the low security task.
When working with
isolated tasks which each have their own heaps, the isolation is obviously true:
the sandboxed task receives a dangling pointer.

While this is a very simple example, we can still ask: given the
obviously safe formulation with separate heaps, can this fact be useful
for showing the safety of a semantics which only has a single heap
(we answer this question positively in Section~\ref{sec:single-heap})?
Furthermore, the na\"ive
semantics violates expectations about how inter-task communication
works on a shared heap: it strongly suggests, for example, that which
types of data may be passed between tasks should be restricted (for
example, in JavaScript, one might only allow JSON objects to be passed
between tasks, rather than general object structures).

@@

When considering tasks, we will instead say that the allocation of a new
stack frame with a raised label corresponds to the creation of a new
isolated task; a sample encoding using our language is shown in
Figure~\ref{fig:toLabeled}.
However, this is not enough: in a stack discipline, one generally
expects stack entries to be processed in order, rather than in parallel
as would be suggested by isolated tasks.  The sequential processing of
the task queue corresponds to a specific \emph{scheduling policy} of
these tasks.  The recasting of the problem to one of scheduling
highlights the essential difference between termination-insensitive and
termination-sensitive non-interference, and we will show how different
scheduling policies give different security guarantees.
